MS Technical Paper: \\ Placement Algorithms for Heterogenous FPGAs

### 2025-03-27 latex/paper/paper.tex
more

    Configurable logic devices have undergone significant evolution over the last four decades.
    The journey began with simple devices like Programmable Logic Array (PLA), which came if the form of logic gates arranged in a "programmable-OR, programmable-AND" plane, used to implement a "sum-of-products" logic equation for each of the outputs in terms of the inputs via programmable fuses. 
    Around the same time came the redundantly named Programmable Array Logic (PAL), which simplified the PLA by fixing the OR gates, resulting in a "fixed-OR, programmable-AND" plane, which sacrificed some flexibility to simplify its manufacture and broaden its applications.

    \begin{figure}
        \centering
        \includegraphics[width=7.0cm]{figures/cpld.png}
        \caption{CPLD architecture with 4 CLBs}
        \label{fig:cpld}
        % https://www.electronics-tutorial.net/Programmable-Logic-Device-Architectures/CPLD/Complex-Programmable-Logic-Device-CPLDs/
    \end{figure}

    Then came the Complex Programmable Logic Device (CPLD), which came as an array of Configurable Logic Blocks (CLBs) stitched together by interconnection wires.
    These CLBs usually took the form of modified PAL blocks, which contained the PAL itself and some macrocells including flip-flops, multiplexers, and tri-state buffers.
    The CPLD was essentially an array of PALs interconnected by a programmable switch matrix and could be programmed via a hardware description language (HDL) like VHDL.

@@

@@

\section{The FPGA Design Flow and Toolchain}
    \begin{figure}
        \centering
        \includegraphics[width=6.5cm]{figures/vivado_dcps.png}
        \caption{Showing how RapidWright can take a design checkpoint at different stages of the Vivado toolchain, make customizations, then feed it back into the toolchain.}
        \label{fig:device_carry_chain_routing}
    \end{figure}
    A typical FPGA design flow goes like the following:
    Engineer design entry (Verilog, VHDL) \\
    Synthesis (Instantiation, substitution, inference, and optimization) \\ 
    Placement (Prepacking, Packing, Global Placement, Legalization, Detailed Placement)\\ 
    Routing \\ 
    Bitstream generation \\

@@

@@

    \subsection{EDIF Package}

>>>>

    Configurable logic devices have undergone significant evolution over the last four decades.
    The journey began with simple devices like Programmable Logic Array (PLA) in the early 1970s, which came if the form of logic gates arranged in a "programmable-OR, programmable-AND" plane, used to implement a "sum-of-products" logic equation for each of the outputs in terms of the inputs via programmable fuses. 
    Around the same time came the redundantly named Programmable Array Logic (PAL), which simplified the PLA by fixing the OR gates, resulting in a "fixed-OR, programmable-AND" plane, which sacrificed some flexibility to simplify its manufacture and broaden its applications.

    \begin{figure}
        \centering
        \includegraphics[width=7.0cm]{figures/cpld.png}
        \caption{CPLD architecture with 4 CLBs (PAL-like blocks)}
        \label{fig:cpld}
        % https://www.electronics-tutorial.net/Programmable-Logic-Device-Architectures/CPLD/Complex-Programmable-Logic-Device-CPLDs/
    \end{figure}

    Later in the same decade came the Complex Programmable Logic Device (CPLD), which came as an array of Configurable Logic Blocks (CLBs).
    These CLBs usually took the form of modified PAL blocks, which contained the PAL itself and some macrocells including flip-flops, multiplexers, and tri-state buffers.
    The CPLD was essentially an array of PALs interconnected by a programmable switch matrix and could be programmed via a hardware description language (HDL) like VHDL.

@@

    The next milestone in the mid-1980s was the advent of homogenous FPGA, which like the CPLD, took the form of a vast array of CLBs. 
    Instead of having a central programmable switch matrix like that of a CPLD, FPGAs adopted an island style architecture, where each CLB, or "island" of logic, is surrounded on all sides by programmable routing resources, resulting in a fabric of programmable logic and routing.

@@

\section{The FPGA Design Flow and Toolchain}

    A typical FPGA design flow goes like the following: 
    Engineer design entry (Verilog, VHDL) \\ 
    Synthesis (Instantiation, substitution, inference, and optimization) \\ 
    Placement (Prepacking, Packing, Global Placement, Legalization, Detailed Placement)\\ 
    Routing \\ 
    Bitstream generation \\

@@

    \begin{figure}
        \centering
        \includegraphics[width=6.5cm]{figures/vivado_dcps.png}
        \caption{Showing how RapidWright can take a design checkpoint at different stages of the Vivado toolchain, make customizations, then feed it back into the toolchain.}
        \label{fig:device_carry_chain_routing}
    \end{figure}

@@

    \subsection{EDIF Package}
    In Vivado, all designs post synthesis have a logical netlist that can be exported in the Electronic Design Interchange Format (EDIF) netlist format.
    The synthesis stage of the FPGA toolchain produces an EDIF file, which is a low-level file containing the technology-mapped netlist.
    The EDIFNetlist is the top level class that contains the netlist and cell libraries. 
    The EDIFNetlist keeps a reference to the top cell which is wrapped in the EDIFDesign class. 
    It also maintains a top cell instance reference that is generated when the file is loaded.
